---
title: "M238_HW1"
author: "Huiyu (Joyce)"
date: "4/8/2019"
output: html_document
---

## Problem 1
(1) Using a convenient statistical package such as R, conduct a simulation study to find the type I error rate of the above group sequential test, the probability under H0 that $\lvert Z_k \rvert >1.96$ for at least one k, k = 1;...; 5.

$$ X_A \sim N(\mu_A, 1); X_B \sim N(\mu_B, 1)$$ 

- Assume $\mu_A = \mu_B = 3$ (under H0), conduct following simulation study. 

```{r}
h <- 1:10000 
for (a in 1:10000) {
X.A <- rnorm(50, 3, 1)
X.B <- rnorm(50, 3, 1)
#sum.A <- sum(X.A)
#sum.B <- sum(X.B)
# divided into k = 5 equally sized groups
data.A <- matrix(X.A, nrow = 5, ncol = 10)
data.B <- matrix(X.B, nrow = 5, ncol = 10)
for (i in 1:5){
  mk <- i*10
  z <- (1/sqrt(2*mk))*(sum(X.A[1:mk]) - sum(X.B[1:mk]))
  if (abs(z) >= 1.96) { 
    h[a] <- 1 
    break
    } 
  h[a] <- 0
}
}
sum(h)/10000
```

- By simulate group sequential test as above, the type I error is close to 0.14 (same as result on lecture slides)

(2) Find the power of the above group sequential test when $ \mu_A - \mu_B$ = 0.1, 0.5, 1.0, 1.5, 2.0

```{r}
m.a <- 5
d <- c(0.1, 0.5, 1.0, 1.5, 2.0)
h <- 1:10000 
power <- c(0,0,0,0,0)
for(b in 1:5){
  m.b <- m.a - d[b]
  for(a in 1:10000) {
    X.A <- rnorm(50, m.a, 1)
    X.B <- rnorm(50, m.b, 1) 
    data.A <- matrix(X.A, nrow = 5, ncol = 10)
    data.B <- matrix(X.B, nrow = 5, ncol = 10)
    for (i in 1:5){
      mk <- i*10
      z <- (1/sqrt(2*mk))*(sum(X.A[1:mk]) - sum(X.B[1:mk]))
      if (abs(z) >= 1.96) {
        h[a] <- 1
        break
      } 
      h[a] <- 0
    }
  }
  power[b] <- sum(h)/10000
}
print(power)
```

- By simulating the data above, the power for $ \mu_A - \mu_B$ with 0.1, 0.5, 1.0, 1.5, 2.0 are about `0.18, 0.77, 0.99, 1.00, 1.00`

(3) Find the expected terminating sample size for the above group sequential test when $ \mu_A - \mu_B$ = 1.0
```{r}
d <- 1.0 
m.a <- 5
m.b <- m.a - d
size <- rep(50, 10000)
sq <- rep(5, 10000)
for(a in 1:10000) {
  X.A <- rnorm(50, m.a, 1)
  X.B <- rnorm(50, m.b, 1) 
  data.A <- matrix(X.A, nrow = 5, ncol = 10)
  data.B <- matrix(X.B, nrow = 5, ncol = 10)
  for (i in 1:5){
    mk <- i*10
    z <- (1/sqrt(2*mk))*(sum(X.A[1:mk]) - sum(X.B[1:mk]))
    if (abs(z) >= 1.96) {
      size[a] <- mk
      sq[a] <- i
      break
    }
    }
}
print(mean(size))
```

- By simulating the data above, expected terminating sample size are arround `15`. 

## Problem 2: 
Repeat problem 1 for the following group sequential test. 

(1) 
```{r}
h <- 1:10000 
for (a in 1:10000) {
  X.A <- rnorm(50, 3, 1)
  X.B <- rnorm(50, 3, 1)
# divided into k = 5 equally sized groups
  for (i in 1:5){
    mk <- i*10
    z <- (1/sqrt(2*mk))*(sum(X.A[1:mk]) - sum(X.B[1:mk]))
    if (i < 5) {
      if (abs(z) >= 3) {
        h[a] <- 1
        break
        }
    }else{
      if (abs(z) >= 1.96){
        h[a] <- 1
        break
      } else {h[a] <- 0}
      } 
  }
}
sum(h)/10000
```

- By simulate group sequential test as above, the type I error is close to `0.05`

(2) 
```{r}
m.a <- 5
d <- c(0.1, 0.5, 1.0, 1.5, 2.0)
h <- 1:10000 
power <- c(0,0,0,0,0)
for(b in 1:5){
  m.b <- m.a - d[b]
  for(a in 1:10000) {
    X.A <- rnorm(50, m.a, 1)
    X.B <- rnorm(50, m.b, 1) 
    for (i in 1:5){
      mk <- i*10
      z <- (1/sqrt(2*mk))*(sum(X.A[1:mk]) - sum(X.B[1:mk]))
      if (i < 5) {
        if (abs(z) >= 3) {
          h[a] <- 1
          break
        }
      }
      else {
        if (abs(z) >= 1.96){
          h[a] <- 1
          break
        } else {h[a] <- 0}
      }
    }
  }
  power[b] <- sum(h)/10000
}
power
```

- By simulating the data above, the power for $ \mu_A - \mu_B$ with 0.1, 0.5, 1.0, 1.5, 2.0 are about `0.07, 0.70, 0.99, 1.00, 1.00`

(3)
```{r}
d <- 1.0 
m.a <- 5
m.b <- m.a - d
size <- rep(50, 10000)
sq <- rep(5, 10000)
for(a in 1:10000) {
  X.A <- rnorm(50, m.a, 1)
  X.B <- rnorm(50, m.b, 1) 
  data.A <- matrix(X.A, nrow = 5, ncol = 10)
  data.B <- matrix(X.B, nrow = 5, ncol = 10)
  for (i in 1:5){
    mk <- i*10
      z <- (1/sqrt(2*mk))*(sum(X.A[1:mk]) - sum(X.B[1:mk]))
      if (i < 5) {
        if (abs(z) >= 3) {
          size[a] <- mk
          break
        }
      } 
  }
}
print(mean(size))
```

- By simulating the data above, expected terminating sample size are arround `24`. 
